---
title: "The accelerated R Course"
subtitle: "Challenges" 
author: "Pol Capdevila"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup1, include=FALSE}
library(learnr)
library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(MASS)
library(fitdistrplus) 
library(dplyr)
#library(gradethis)
#gradethis::gradethis_setup()
#tutorial_options(exercise.checker = gradethis::gradethis_exercise_checker(),
#                 exercise.error.check.code=gradethis::grade_code())

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Introduction

In today's assignment you are going to put in practice all the skills you have learned in the previous lessons. You will have a number of tasks to complete in 1h 30' by working in groups of __five__ people.

Each task contains a number of exercises that you need to solve before completing the overall task. Some of the tasks contain data that you can download [in this link](https://github.com/PolCap/Teaching/tree/master/Bristol/R%20Course/data) (the links will also be provided in each exercise) and run in your computer. You will also have some space to run the code within the  Once you have completed a task, please tell us and we will come to check whether it is correct or not. 

---

<center>__The first group to complete each task will receive a little reward.__</center>

---


## Task 1

It is time for you to put a bit on practice your programming skills with R. In this task you are going to solve some specific exercises 

### Exercice 1.1 

* Write a function called `help` that when executed prints a message ("Help").

```{r exercise1_1, exercise = TRUE,exercise.lines=10}





```

```{r exercise1_1-hint}
function(x){
  #your function
  }
```

### Exercice 1.2 

* Create a function that given a vector `random_vector` and an `integer` will return how many times that `integer` appears inside the vector.

```{r exercise1_2, exercise = TRUE,exercise.lines=10}
random_vector <- c(8,9,9,8,5,1,8,5,6,2,3,5,9,9,8)


```

### Exercice 1.3 

* Create a function that when given a `data frame`, and a specific `number` or `character`, will return the `data frame` with that character or number changed to `NA`

```{r exercise1_3, exercise = TRUE,exercise.lines=10}





```

### Exercice 1.4 

* Use the commands `if` and `else` to return the even numbers only from the given vector. If you are unsure how to use `if else`, you can check [this link](https://www.datamentor.io/r-programming/if-else-statement/)

```{r exercise1_4, exercise = TRUE, ,exercise.lines=10}
odds_even <- c(0,2,3,5,6,8,9)





```
<div id="exercise1_4-hint">
**Hint:** To identify pair numbers you can use `x %% 2 == 0`.
</div>

## Task 2

In this second task you will work with the [Living Planet database](https://github.com/PolCap/Teaching/tree/master/Bristol/R%20Course/data), a database containing abundance time-series for wild vertebrate populations worldwide.  

```{r data, message=FALSE,echo=FALSE, warning=FALSE}
lpd <- read_csv("data/LivingPlanetIndex.csv")

head(lpd)
```

As you can see, the database has five columns with metadata and from column 6 to 74 we have abundance data from the year 1950 to 2018. In the metadata we have: 

- __ID__: The unique identification for each population. 
- __Species__: The name of the species of that population. 
- __Class__: The name of the taxonomic class of the species. 
- __Country__: Country where the population was studied. 
- __System__: Whether it was studied in a Marine, Freshwater or Terrestrial system. 

Now that you understand the data a bit more, is time for you to reply to each of the following exercises. You can use the panels provided to run the code.  


### Exercice 2.1

```{r quiz2_1}
quiz(
  question("What is the earliest year that the UK started to monitor the abundance of a population of vertebrates (according to the living planet index)?",
    answer("1903"),
    answer("1967"),
    answer("1950", correct = TRUE),
    answer("1953")
  ))
```


```{r exercise2_1, exercise=TRUE,exercise.lines=10}
lpd 


```


###  Exercise 2.2

```{r quiz2_2}
quiz(
  question("What is the most recent year that the UK started monitoring a new population (according to the living planet index)?",
    answer("2015", correct = TRUE),
    answer("2018"),
    answer("2016"),
    answer("2017")
  )
)
```

```{r exercise2_2, exercise=TRUE, exercise.lines=10}
lpd 

```


### Exercice 2.3

```{r quiz2_3}
quiz(
  question("Which country has the the longest time-series?",
    answer("United States"),
    answer("Canada"),
    answer("South Africa", correct = TRUE),
    answer("Spain", correct = TRUE)
  ))
```

```{r exercise2_3, exercise=TRUE,exercise.lines=10}


```

## Task 3 

In this task you will use the `foxes.csv` data set which contains data on urban fox groups in England. [Link here](https://github.com/PolCap/Teaching/tree/master/Bristol/R%20Course/data). 

```{r foxesdata, eval=T,warning=FALSE}
foxes <- read_csv("data/foxes.csv")
foxes

```


As you probably already know, the urban fox (_Vulpes vulpes_) is a successful exploiter of human habitat. Since urban foxes move in packs and defend territories, data on habitat quality and population density is also included. The data frame has five columns:

- __group__: Number of the social group the individual fox belongs to
- __avgfood__: The average amount of food available in the territory
- __groupsize__: The number of foxes in the social group
- __area__: Size of the territory
- __weight__: Body weight of the individual fox

In this task you will have to find which are the best predictors for fox body weight. 

### Exercise 3.1 

```{r quiz3_1}
quiz(
  question("Which is the best distribution for `weight` data?",
    answer("Poisson"),
    answer("Binomial"),
    answer("Gaussian", correct = TRUE),
    answer("Negative binomial")
  ))
```

* __Note__: You will have to justify your choice. 

```{r exercise3_1, exercise=TRUE, exercise.lines=20}
foxes

```

### Exercise 3.2 

* Fit a model which will allow you to determine which of the four predictor variables has the strongest influence on the body size of foxes. 

```{r exercise3_2, exercise = TRUE,exercise.lines=20}
foxes

```


## Task 4

__This is your final task! You are nearly there.__ 


In this task you will use the `starwars` data set, [link here](https://github.com/PolCap/Teaching/tree/master/Bristol/R%20Course/data). This data set contains morphological characteristics of the different characters appearing in the Starwars movies (up to the 7th film). We want to calculate the average height of men and women for each of the homeworlds. The wookies have produced some code that works "fine", but we would like to improve it in order to be more efficient.   

### Exercise 4.1 

* Re-write the following code using lapply():


```{r exercise4_1, exercise=T, exercise.lines=20}

average_height <- list()


for(i in starwars$homeworld){
  for(j in starwars$gender){
    substar <- subset(starwars, homeworld==i&gender==j)
    average <- mean(substar$height)
  }
}

```

### Exercise 4.2 

* Achieve the same result using the tidyverse functions:


```{r exercise4_2, exercise=T, exercise.lines=20}

average_height <- list()
for(i in starwars$homeworld){
  for(j in starwars$gender){
    substar <- subset(starwars, homeworld==i&gender==j)
    average <- mean(substar$height)
  }
}

```


